---
title: "Homework"
output: 
  html_document:
    theme: readable
    # toc: true
    # toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(eval = FALSE)
```
Given the following graph:

```{r}
title <- 'Fuel Economy of Popular Cars'
legend.title <- 'Type of Car'

ggplot(mpg, aes(displ, hwy, color = class)) + 
  geom_point()
```

**Which are ways that you can change the legend titles? Select all that apply.**

  a. `scale_color_discrete(name = legend.title)`
  b. `theme(legend.title = legend.title)`
  c. `theme(legend.text = element_text(title = legend.title))`
  d. `labs(color = legend.title)`

**Which are ways that you can add a title to your graph? Select all that apply.**

  a. `labs(title = title)`
  b. `ggtitle(title)`
  c. `annotate('text', label = title, x = min(mpg$displ) + 3.5, y = max(mpg$hwy), size = 4)`
  d. `theme(plot.title = element_text(title))`

***

**Create a scatterplot exploring select breakfast cereals[^1]. What is the relationship between cereal ratings and grams of sugar?**  

[^1]: Data Source: https://www.kaggle.com/crawford/80-cereals, gathered and cleaned up by Petra Isenberg, Pierre Dragicevic and Yvonne Jansen. Original source can be found [here](https://perso.telecom-paristech.fr/eagan/class/igr204/datasets)  


  1. Download the [cereals dataset](https://github.com/psrc/intro-ggplot2/raw/main/data/cereal.xlsx), and read into R. Metadata can be found [here](https://www.kaggle.com/crawford/80-cereals)
  2. Create a new column `sugars_per_oz` that calculates grams of sugars per ounce. 
  3. Create a scatterplot using `geom_point()` with `sugars_per_oz` on the x-axis and rating on the y-axis
  4. Map the manufacturer of the cereal to color and cereal type to shape 
  5. Change the shapes to anything but the default ones
      - To find out what shape options are available run the following
      
```{r}
df_shapes <- data.frame(shape = 0:24)
ggplot(df_shapes, aes(0, 0, shape = shape)) +
  geom_point(aes(shape = shape), size = 5, fill = 'red') +
  scale_shape_identity() +
  facet_wrap(~shape) +
  theme_void()
```

**Notice anything interesting with the sugar column in the dataset?**  

  6. What is the range of the sugars per ounce column?
  7. Which cereal(s) contains the greatest amount of sugars per ounce? Cereals with 0 grams? Less than 0?
  
**On which shelf can you find the cereal with the highest rating?**  

  8. Facet wrap by display shelf (1, 2, or 3, counting from the floor)
  9. Rename x & y axis labels
  10. Rename legend titles
  11. Add a label using `geom_text()` for the cereal with the highest rating
  12. Things might look a bit squished. Try reducing the size of the label and the legend labels

***

**Create a bar graph of cities/towns with greatest nominal growth between 2010 and 2020**  

Using `ofm_april1_population_final_tidied.xlsx`...

  1. Subset for cities/towns and where year is 2010 and 2020
  2. Cast the data using the `dcast()` function from the `reshape2` package: 
      - `reshape2::dcast(<your data frame>, County + Jurisdiction ~ paste0("Year_", Year_chr), value.var = "Estimate")`
  3. Create a new column containing the difference
  4. Sort the data frame so that the difference column is in descending order
  5. Take the top 10 rows (use `head()`)
  6. Plot where cities/towns are on the x-axis and the y-axis displays total population estimate
      - Add thousands separator to the Y axis column
      - Angle the cities/town axis labels and reduce the font size
      - Change the y-axis title to something more descriptive
      - Add a title and source caption
  7. What order are the cities/town in? Can you reorder the cities/towns based on the value of the difference column?
      - convert the `Jurisdiction` column to a factor
      - use `reorder()`


