---
title: "Using Fonts with ggplot2 in Windows"
output: 
  html_document:
    theme: readable
    toc: true
    toc_float: true
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(eval = FALSE)
```

Fonts are tricky to work with in R. Unfortunately R and ggplot2 don't recognize fonts outside of their standard 3 without a little work. Below are some one-time steps to allow R to recognize different fonts for bitmap and screen output (e.g. RStudio's Plot Pane, .jpg, .png) and PDFs. 

The instructions below are adapted from the ones in the [extrafont vignette](https://cran.r-project.org/web/packages/extrafont/README.html) for a Windows machine. I've made some minor changes and omitted steps (e.g. GhostScript installation) in order for it to work on my computer. 

Take the following code snippets and paste/run in the console of your RStudio IDE.

# Windows Bitmap and Screen Output

If you want to use fonts other than the basic ones R recognizes and want it to render in the RStudio Plot Pane and in an exported bitmap file like .jpg, .png:

1. Install the `extrafont` package. This package helps identify the fonts on your machine and handles a lot of the details for you.
```{r}
install.packages("extrafont")
```

2. Load the `extrafont` library
```{r}
library(extrafont)
```

3. Run the font import function so that R can find and save information about fonts installed on your system. This step may take a while.
```{r}
font_import()
```

4. Register the fonts for Windows
```{r}
loadfonts("win")
```

## Test
Let's test it! See if this code will produce a plot in your plot pane
```{r}
my.plot <- ggplot(mtcars, aes(x=wt, y=mpg)) + 
  geom_point() +
  ggtitle("Fuel Efficiency of 32 Cars") +
  xlab("Weight (x1000 lb)") + 
  ylab("Miles per Gallon") +
  theme(text = element_text(size=16, family="Impact"))

print(my.plot)
```

Now see if you can save it as a .jpg. Change the file path to a folder that works for you.
```{r}
ggsave("T:/2020November/christy/my_plot.jpg", plot = my.plot, width = 4, height = 4)
```


# PDFs

After completing the steps in Windows Bitmap Output, you also have the option of exporting your plot with custom font to PDF but it will require a special argument `device = cairo_pdf` when you use `ggsave()`. 


```{r}
ggsave("T:/2020November/christy/my_plot.pdf", plot = my.plot,  width=4, height=4, device = cairo_pdf)

```

# Font choices

To find the fonts available for you to use after setup:
```{r}
# Vector of font family names
fonts()

# Show entire table
font.df <- fonttable()
View(font.df)

```


